---
title: "PML_Project_v01"
author: "Andy Williams"
date: "12/08/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Machine Learning

This rmd file is a submission for the Coursera Course "Practical Machine
Learning"

### Step 1

Load in the required libraries and data

** Datasource: Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. 
Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th
International Conference in Cooperation with SIGCHI (Augmented Human '13) .
Stuttgart, Germany: ACM SIGCHI, 2013

Read more: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

```{r, message=FALSE,warning=FALSE}
set.seed(1234)

library(caret)
library(ggplot2)
library(rattle)
library(dplyr)

#setwd("PML_Project")

df <- read.csv("pml-training.csv")
df_test <- read.csv("pml-testing.csv")

```

### Assess spread of data for each user vs each class

For each user 10 repetitions of the Unilateral Dumbbell Biceps Curl were done in
five different fashions: exactly according to the specification (Class A), 
throwing the elbows to the front (Class B), lifting the dumbbell only halfway 
(Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to
the front (Class E).

```{r}
table(df$user_name,df$classe)
```
The data shows that each participant completed all intended exercises. There is 
a potential bias for class A which is the largest exercise assessed.

### Clean data

Remove any columns with 100% NAs or non numerical values

```{r}

print(paste("Columns before cleaning = ",ncol(df)))

df_clean <- df[ , colSums(is.na(df)) == 0]

print(paste("Columns after cleaning for NA = ",ncol(df_clean)))

df_labels <- df_clean$classe

df_clean <- select_if(df_clean, is.numeric)

df_clean <- cbind(df_clean,df_labels)
df_clean$classe <- df_clean$df_labels
df_clean <- select(df_clean,-c("df_labels","X"))

print(paste("Columns after cleaning for non numeric columns = ",ncol(df_clean)))

```

## Feature Plot

Assess data as a feature plot for example of the belt wearable 

```{r}
featurePlot(y=as.factor(df_clean$classe), x=df_clean[,c(2:14)] ,data=df_clean, plot = "pairs")
```

### Cross Validate and train

As the data contains sequenced data similar to time series it was decided to use
repeated KFolds to chunk up the data

LDA and SVM models were assessed.

```{r}
set.seed(54321)
train_control <- trainControl(method = "repeatedcv", number = 10, repeats = 5)
modelLDA <- train(classe ~ ., data = df_clean, method = "lda", trControl = train_control)
modelSVM <- train(classe ~ ., data = df_clean, method = "svmLinear", trControl = train_control)
```

### Assess accuracy of each model

```{r}

print(paste("Accuracy of LDA method:",modelLDA$results$Accuracy))
print(paste("Accuracy of SVM method:",modelSVM$results$Accuracy))

```
### Selection of method

The SVM model was selected to predict the test results

### Predicted class's

```{r}

predSVM <- predict(modelSVM, df_test[,-160])
print(predSVM)

```

